---
title: "02a_beeReproDev_DataExploration"
author: "Jessie Dodge"
date: "2024-05-23"
output:
  word_document: default
  pdf_document: 
    latex_engine: xelatex
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE) 
```

# Library
```{r, include = FALSE}
source(file="01i_beeReproDev_MasterData.R")
```

# Data org
```{r}
# To add labels for sites sampled both years
data<-data %>% group_by(Site) %>%
  mutate(Year2 = n_distinct(Year)) %>% 
  #mutate(Year3 = as.factor(ifelse(Year2 == 2, "B", "")))
  mutate(Year3 = as.factor(ifelse(Year2 == 2, Site, "")))

# turn SiteYr to rownames
data<-data %>% column_to_rownames(var = "SiteYr")

# re org and select columns you want to keep
data2<-data[,c(1:3, 14:42, 46:49, 53, 63:64, 67:75)]
# Count/proportion data 
colSums(data2[,c(5:20, 43:48)] > 0) # Crops are only in 3 sites, and Barren are only in 2 sites
data2<-data2[,-c(43,45)] # omit crops and barren

# omit site with no provisioned cells - zeros omitted (zo)
zo.data<-filter(data2, nTotProvised > 0)

# Data subsets
# for count/weight data
## all
or.data <- filter(zo.data, oligRatio > 0)
w.data <- filter(zo.data, avgOligMass_mg > 0)
fw.data <- filter(zo.data, avgOligFMass_mg > 0)
mw.data <- filter(zo.data, avgOligMMass_mg > 0)
lw.data <- filter(zo.data, avgOligLarvMass_mg > 0)

# remove T. stan outlier
zo.data.or <- filter(zo.data, nTstanCells < 10)
fw.data.or <- filter(zo.data.or, avgOligFMass_mg > 0)
fw.data.or2 <- filter(zo.data.or, avgOligFMass_mg > 0 & Year == 2021)
mw.data.or <- filter(zo.data.or, avgOligMMass_mg > 0)
lw.data.or <- filter(zo.data.or, avgOligLarvMass_mg > 0)
or.data.or <- filter(or.data, nTstanCells < 10 & Year == 2021)
```
# Questions
Q1) How does landscape and climate variation affect bee parasitism?
  * Proposed analyses: MLRs 
    ** Parasite abundance ~ Year + Hab + Landscape + Climate + Floral
Q2) How does parasitism abundance affect bee fitness traits?
  * Proposed analyses: SLRs 
    ** bee % ~ parasite %
    ** bee abundance ~ parasite abundance
    ** bee mass ~ parasite abundance (2021)
    ** F% ~ parasite abundance (2021)

# Data distribution 
## O. lignaria
### All sites
```{r, include=FALSE, echo=FALSE}
# All sites
## Ratios
hist(data$oligRatio) # sort of normal? After ignoring zeros
hist(sqrt(data$oligRatio)) # worse
hist(log(data$oligRatio +1, base=10)) #  still zero heavy, but a bit better
hist(1/(data$oligRatio)) # right skewed 

## counts
hist(data$nTotProvised) # skewed, but not quite poisson (not zero heavy)
hist(sqrt(data$nTotProvised)) # more normal
hist(log(data$nTotProvised+1,base=10)) # a bit left skewed but better
hist(log(data$nTotProvised+1)) # VERY left skewed
hist(1/(data$nTotProvised)) # worse

### Ratio by year
ggplot(data, aes(x = oligRatio)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 25, alpha = 0.4) 
  # zero heavy - especially in 2021

### Counts by year
ggplot(data, aes(x = nTotProvised)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 25, alpha = 0.4)
  # zero heavy - especially in 2021
```

### O.lig ratio > 0 Sites
```{r}
## Ratio
hist(or.data$oligRatio) # pretty normal
hist(or.data.or$oligRatio) # right skewed
hist(fw.data.or2$oligRatio) # pretty normal

## Counts
hist(or.data$nOligGrowth) # right skewed
hist(sqrt(or.data$nOligGrowth)) # more normal
hist(log(or.data$nOligGrowth +1, base=10)) # bi-modal

### Ratio by year
ggplot(or.data, aes(x = oligRatio)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 20, alpha = 0.4)
  # lower ratios in 2021

### Counts by year
ggplot(or.data, aes(x = nTotProvised)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 20, alpha = 0.4)
  # years are pretty similar
```

### Provisioned Sites
```{r}
hist(zo.data$nOligGrowth) # right skewed
hist(sqrt(zo.data$nOligGrowth)) # pretty normal
hist(log(zo.data$nOligGrowth+1, base=10)) # left skewed

### Counts by year
ggplot(zo.data, aes(x = nTotProvised)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 20, alpha = 0.4) 
  # years are pretty similar
```

### O.lig weights 
```{r}
# female
hist(fw.data$avgOligFMass_mg) # right skewed
hist(sqrt(fw.data$avgOligFMass_mg)) # more normal
hist(fw.data.or$avgOligFMass_mg) # Gaussian
hist(fw.data.or2$avgOligFMass_mg) # sort of Gaussian
hist(log(fw.data.or2$avgOligFMass_mg, base=10)) # better

# Male
hist(mw.data$avgOligMMass_mg) # fairly normal
hist(mw.data.or$avgOligMMass_mg) # fairly normal

# Larvae
hist(lw.data$avgOligLarvMass_mg) # normal
hist(lw.data.or$avgOligLarvMass_mg) # normal

### Counts by year
ggplot(fw.data, aes(x = avgOligFMass_mg)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 20, alpha = 0.4) 
  # only 5 females in 2020, but within the dist. of 2021 weights

ggplot(mw.data, aes(x = avgOligMMass_mg)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 20, alpha = 0.4) 
  # 2021 is normally distributed. Not many males in 2020
```

## T.stansburyi
Provisioned Sites - would only be present if there were provisioned cells
```{r}
# with outlier
hist(zo.data$nTstanCells, breaks = 20) # poisson dist - HUGE outlier
hist(sqrt(zo.data$nTstanCells)) # still skewed
hist(log(zo.data$nTstanCells+1, base=10)) # more poisson shaped now

hist(zo.data$tstanRatio) # poisson dist - outlier not as problematic
hist(sqrt(zo.data$tstanRatio)) # 0 heavy
hist(log(zo.data$tstanRatio+1, base=10)) # worse

# without outlier
hist(zo.data.or$nTstanCells, breaks = 10) # very zero heavy
hist(sqrt(zo.data.or$nTstanCells)) # normalish
hist(log(zo.data.or$nTstanCells+1,base=10)) # same as above

hist(zo.data.or$tstanRatio) # very zero heavy
hist(sqrt(zo.data.or$tstanRatio)) # sort of normal
hist(log(zo.data.or$tstanRatio+1, base=10)) # right skewed

### Counts by year
ggplot(zo.data, aes(x = nTstanCells)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 20, alpha = 0.4) 
  # years are pretty similar
ggplot(zo.data.or, aes(x = nTstanCells)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 10, alpha = 0.4) 
  # years are pretty similar
```

## Climate variables
```{r}
hist(zo.data$HLI)# zero heavy
hist(zo.data$Ann_ppt_mm) # left skewed
hist(zo.data$Ann_tmax_C) # left skewed
hist(zo.data$Ann_tmean_C) # normalish
hist(zo.data$Ann_tmin_C) # slightly right skewed
hist(zo.data$bi_ppt_mm) # bi-modal
hist(zo.data$bi_tmin_C) # normalish?
hist(zo.data$bi_tmean_C) # normalish
hist(zo.data$bi_tmax_C) # normalish
hist(zo.data$qrt_ppt_mm) # bi-modal
hist(zo.data$qrt_tmin_C) # normalish
hist(zo.data$qrt_tmean_C) # normalish
hist(zo.data$qrt_tmax_C) # normalish
hist(zo.data$DD_accumulated) # bi-modal

ggplot(zo.data, aes(x = DD_accumulated)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 10, alpha = 0.4) 
# very different by years
```

## Floral variables
```{r}
# stemrich
hist(zo.data$stemrich) # right skewed
hist(sqrt(zo.data$stemrich)) # more normal
hist(log(zo.data$stemrich, base=10)) # normal

ggplot(zo.data, aes(x = stemrich)) + 
  geom_histogram(aes(color = Year, fill = Year), position = "identity", bins = 10, alpha = 0.4) 
# pretty different by year
```

## Landscape variables
Provisioned Sites
```{r}
# HLI
hist(zo.data$HLI) # left skewed - hot and dry sites?
# don't know how to transform
hist(zo.data$HLI^2) # left skewed
hist(abs(zo.data$HLI)) # zero heavy
hist(sqrt(abs(zo.data$HLI))) # bimodal
hist(log(abs(zo.data$HLI)+1,base=10))
hist(log(abs(zo.data$HLI)+1)) # still left skewed, but not as bad
hist(1/(abs(zo.data$HLI)))

# TCC
hist(zo.data$meanTreeCC) # zero heavy - burned sites
hist(sqrt(zo.data$meanTreeCC)) # still zero heavy
hist(log(zo.data$meanTreeCC, base = 10)) # Left skewed
hist(1/(zo.data$meanTreeCC)) # right skewed

# EVT
hist(zo.data$EVT_abund) # normalish
hist(zo.data$EVT_H) # sort of right skewed
hist(zo.data$EVT_rich) # normalish, with many breaks
hist(zo.data$EVT_J) # normalish
hist(zo.data$Barren) # VERY right skewed
hist(zo.data$Coniferous_Forest) # VERY left skewed
hist(zo.data$Crops) # VERY right skewed
hist(zo.data$Deciduous_Forest) # VERY right skewed
hist(zo.data$Rangeland) # VERY right skewed
hist(zo.data$WUI) # VERY right skewed
```

# Normality
## O. lignaria
### All sites
```{r}
# Ratios
qqnorm(data$oligRatio) # not quite normal
shapiro.test(data$oligRatio) # technically not sig (p = 0.07)

## O.lig ratio > 0 sites
qqnorm(or.data$oligRatio) # fairly normal
shapiro.test(or.data$oligRatio) # not sig
qqnorm(sqrt(or.data$oligRatio)) # more normal

# counts
qqnorm(data$nTotProvised) # not quite normal
shapiro.test(data$nTotProvised) # sig - not normal
qqnorm(sqrt(data$nTotProvised)) # pretty normal
shapiro.test((sqrt(data$nTotProvised))) # not sig - normal

# weights
## 2020
qqnorm(fw.data$avgOligFMass_mg) # too few obs
qqnorm(mw.data$avgOligMMass_mg) # normal
qqnorm(lw.data$avgOligLarvMass_mg) # normal
```
### Provisioned sites
```{r}
# Ratios
qqnorm(zo.data$oligRatio) # pretty normal
shapiro.test(zo.data$oligRatio) # not sig

## O.lig ratio > 0 sites
qqnorm(zo.data$oligRatio) # fairly normal
shapiro.test(zo.data$oligRatio) # not sig

# counts
qqnorm(zo.data$nTotProvised) # not quite normal
shapiro.test(zo.data$nTotProvised) # sig - not normal
qqnorm(sqrt(zo.data$nTotProvised)) # sort of normal
shapiro.test((sqrt(zo.data$nTotProvised))) # not sig - normal
```
### O.lig ratio > 0 Sites
```{r}
# Ratios
qqnorm(or.data$oligRatio) # fairly normal
qqnorm(sqrt(or.data$oligRatio)) # more normal
qqnorm(log(or.data$oligRatio+1, base=10)) # sort of normal

## Counts
qqnorm(or.data$nOligGrowth) # wonky
qqnorm(sqrt(or.data$nOligGrowth)) # sort of normal?
shapiro.test(sqrt(or.data$nOligGrowth)) # not sig
qqnorm(log(or.data$nOligGrowth +1, base=10)) # eehhhh
```

## T. stansburyi data
### All sites
```{r}
## with outlier
qqnorm(data$nTstanCells) # not normal - outlier and zero heavy
qqnorm(sqrt(data$nTstanCells)) # still zero heavy
qqnorm(log(data$nTstanCells+1,base = 10)) # still zero heavy

qqnorm(data$tstanRatio) # not normal - outlier and zero heavy
qqnorm(sqrt(data$tstanRatio)) # still zero heavy
qqnorm(log(data$tstanRatio+1,base = 10)) # still zero heavy

## without outlier
qqnorm(zo.data.or$nTstanCells) # not normal - outlier and zero heavy
qqnorm(sqrt(zo.data.or$nTstanCells)) # still zero heavy
qqnorm(log(zo.data.or$nTstanCells+1,base = 10)) # still zero heavy

qqnorm(zo.data.or$tstanRatio) # not normal - outlier and zero heavy
qqnorm(sqrt(zo.data.or$tstanRatio)) # still zero heavy
qqnorm(log(zo.data.or$tstanRatio+1,base = 10)) # still zero heavy
```

### Provisioned sites
```{r}
qqnorm(zo.data$nTstanCells) # huge outlier
qqnorm(sqrt(zo.data$nTstanCells)) # huge outlier
qqnorm(log(zo.data$nTstanCells+1,base = 10)) # still zero heavy

qqnorm(zo.data$tstanRatio) # not normal - outlier and zero heavy
qqnorm(sqrt(zo.data$tstanRatio)) # still zero heavy
qqnorm(log(zo.data$tstanRatio+1,base = 10)) # still zero heavy
shapiro.test(log(zo.data$tstanRatio+1,base = 10)) # sig
```
### O.lig ratio > 0 Sites
```{r}
# Ratios
qqnorm(or.data$nTstanCells) # huge outlier
qqnorm(sqrt(or.data$nTstanCells)) # huge outlier
qqnorm(log(or.data$nTstanCells+1,base = 10)) # still zero heavy

qqnorm(or.data$tstanRatio) # not normal - outlier and zero heavy
shapiro.test(or.data$tstanRatio) #  sig
qqnorm(sqrt(or.data$tstanRatio)) # still zero heavy
shapiro.test(sqrt(or.data$tstanRatio)) # not sig
qqnorm(log(or.data$tstanRatio+1,base = 10)) # still zero heavy
shapiro.test(log(or.data$tstanRatio+1,base = 10)) # sig

## Counts
qqnorm(or.data$nTstanCells) # wonky
qqnorm(sqrt(or.data$nTstanCells)) # wonky
shapiro.test(sqrt(or.data$nTstanCells)) # sig
qqnorm(log(or.data$nTstanCells +1, base=10)) # zero heavy, then normal?
shapiro.test(log(or.data$nTstanCells+1, base=10)) # p = 0.05.
```

## Climate variables
```{r}
qqnorm(zo.data$HLI)# zero heavy
qqnorm(zo.data$Ann_ppt_mm) # left skewed
qqnorm(zo.data$Ann_tmax_C) # left skewed
qqnorm(zo.data$Ann_tmean_C) # normalish
qqnorm(zo.data$Ann_tmin_C) # slightly right skewed
qqnorm(zo.data$bi_ppt_mm) # bi-modal
qqnorm(zo.data$bi_tmin_C) # normalish?
qqnorm(zo.data$bi_tmean_C) # normalish
qqnorm(zo.data$bi_tmax_C) # normalish
qqnorm(zo.data$qrt_ppt_mm) # bi-modal
qqnorm(zo.data$qrt_tmin_C) # normalish
qqnorm(zo.data$qrt_tmean_C) # normalish
qqnorm(zo.data$qrt_tmax_C) # normalish
qqnorm(zo.data$DD_accumulated) # bi-modal
```

## Floral & Tree variables
```{r}
# stemrich
qqnorm(zo.data$stemrich) # fairly normal

# TCC
qqnorm(zo.data$meanTreeCC) # zero heavy - burned sites
qqnorm(sqrt(zo.data$meanTreeCC)) # still zero heavy
qqnorm(log(zo.data$meanTreeCC, base = 10)) # Left skewed
qqnorm(1/(zo.data$meanTreeCC)) # right skewed
```

## Landscape variables
Provisioned Sites
```{r}
# HLI
qqnorm(zo.data$HLI) # left skewed - hot and dry sites?
# don't know how to transform
qqnorm(zo.data$HLI^2) # left skewed
qqnorm(abs(zo.data$HLI)) # zero heavy
qqnorm(sqrt(abs(zo.data$HLI))) # bimodal
qqnorm(log(abs(zo.data$HLI)+1,base=10))
qqnorm(log(abs(zo.data$HLI)+1)) # still left skewed, but not as bad
qqnorm(1/(abs(zo.data$HLI)))

# EVT
qqnorm(zo.data$EVT_abund) # normalish
qqnorm(zo.data$EVT_H) # sort of right skewed
qqnorm(zo.data$EVT_rich) # normalish, with many breaks
qqnorm(zo.data$EVT_J) # normalish
qqnorm(zo.data$Barren) # VERY right skewed
qqnorm(zo.data$Coniferous_Forest) # VERY left skewed
qqnorm(zo.data$Crops) # VERY right skewed
qqnorm(zo.data$Deciduous_Forest) # VERY right skewed
qqnorm(zo.data$Rangeland) # VERY right skewed
qqnorm(zo.data$WUI) # VERY right skewed
```

# Boxplots
```{r}
boxplot(data$HLI)
boxplot(data$nTstanCells) # HUGE outlier
boxplot(data$nParasite)
boxplot(data$avgOligFMass_mg)
boxplot(data$avgOligMMass_mg)
boxplot(data$avgOligFMass_mg)
boxplot(data$avgOligLarvMass_mg) # outlier, but probably more useful than pupae mass
boxplot(data$oligRatio) # outliers (no bees)
boxplot(data$tstanRatio) # outlier
boxplot(data$monoRatio)
boxplot(data$Dead_BA_ha)
boxplot(data$Live_BA_ha)
boxplot(data$meanTreeCC)
boxplot(data$stemdiv)
boxplot(data$stemrich)
boxplot(data$stemabun)
```

# Identifying outliers
```{r}
# Outliers:
## nTstan/TstanRatio

ggplot(data=data, aes(x=Habitat, y=nTstanCells, label = Site, color=Habitat))+
  geom_boxplot()+
  geom_text(check_overlap = TRUE, position=position_jitter(width=0.01))+
  theme(legend.position="right") # H18 - TON of T. stan

ggplot(data=zo.data.or, aes(x=WUI, y=nTstanCells, label = Site))+
  geom_boxplot()+
  geom_text(check_overlap = TRUE, position=position_jitter(width=0.01))+
  theme(legend.position="right") # H18 - TON of T. stan

ggplot(data=data, aes(x=Habitat, y=tstanRatio, label = Site, color=Habitat))+
  geom_boxplot()+
  geom_text(check_overlap = TRUE, position=position_jitter(width=0.01))+
  theme(legend.position="right") # H18 not as extreme

ggplot(data=data, aes(x=Habitat, y=avgOligLarvMass_mg, label = Site, color=Habitat))+
  geom_boxplot()+
  geom_text(check_overlap = TRUE, position=position_jitter(width=0.01))+
  theme(legend.position="right") # 2 sites with no larvae (H03 & T02)

ggplot(data=data, aes(x=Habitat, y=oligRatio, label = Site, color=Habitat))+
  geom_boxplot()+
  geom_text(check_overlap = TRUE, position=position_jitter(width=0.01))+
  theme(legend.position="right") # H03 - no Olig
```

# Zero inflation?
## All sites
```{r}
# count data (that you might be using)
100*sum(data$nTotProvised == 0)/nrow(data) # 17.9%
100*sum(data$nTstanCells == 0)/nrow(data) # 38.5% - High!

# ratio data
100*sum(data$oligRatio == 0)/nrow(data) # 20.5%
100*sum(data$oligFRatio == 0)/nrow(data) # 46.1% - 2020 probably skewing data
100*sum(data$oligFRatio == 0 & data$Year == 2021)/nrow(data) # 20.5%  - better
100*sum(data$tstanRatio == 0)/nrow(data) # 38.5% - High
```

## Provisioned sites
```{r}
# count data 
100*sum(zo.data$nTstanCells == 0)/nrow(zo.data) # 25 % 
### not too bad, as far as zero inflation, goes

# ratio data
100*sum(zo.data$oligRatio == 0)/nrow(zo.data) # 3.1%
100*sum(zo.data$oligFRatio == 0)/nrow(zo.data) # 34.4% - 2020 probably skewing data
100*sum(zo.data$oligFRatio == 0 & zo.data$Year == 2021)/nrow(zo.data) # 6.2%  - Way better!
100*sum(zo.data$tstanRatio == 0)/nrow(zo.data) # 25%
```
Not too bad, as far as zero inflation goes

# Difference between years
## O. lignaria
### Plots
```{r}
## Count
boxplot(nTotProvised ~ Year, data = zo.data) # not that diff
boxplot(nOligGrowth ~ Year, data = zo.data) # no diff
boxplot(nOligLarve ~ Year, data=zo.data) # much more in 2020
boxplot(nOligF ~ Year, data = zo.data) # much more in 2021
boxplot(nOligM ~ Year, data = zo.data) # much more in 2021

## Mass
boxplot(avgOligMass_mg ~ Year, data = zo.data) # higher in 2021
boxplot(avgOligFMass_mg ~ Year, data = zo.data) # higher in 2021
boxplot(avgOligMMass_mg ~ Year, data = zo.data) # higher in 2021

## Ratio
boxplot(oligRatio ~ Year, data = zo.data) # similar
boxplot(monoRatio ~ Year, data = zo.data) # similar
```
Mass is probably significantly different - different collection methods

### Analyses
```{r}
## Count
bmod1<-aov(nTotProvised ~ Year, data = zo.data) 
bmod1<-tidy(bmod1) # no diff
kable(bmod1, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod2<-aov(nOligGrowth ~ Year, data = zo.data)
bmod2<-tidy(bmod2) # no diff
kable(bmod2, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod3<-aov(nOligLarve ~ Year, data=zo.data) 
bmod3<-tidy(bmod3) # very diff
kable(bmod3, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod4<-aov(nOligF ~ Year, data = zo.data) 
bmod4<-tidy(bmod4) # very diff
kable(bmod4, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod5<-aov(nOligM ~ Year, data = zo.data) 
bmod5<-tidy(bmod5) # no diff
kable(bmod5, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

## Mass
bmod6<-aov(avgOligMass_mg ~ Year, data = zo.data) 
bmod6<-tidy(bmod6) # diff
kable(bmod6, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod7<-aov(avgOligFMass_mg ~ Year, data = zo.data) 
bmod7<-tidy(bmod7) # diff
kable(bmod7, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod8<-aov(avgOligMMass_mg ~ Year, data = zo.data) 
bmod8<-tidy(bmod8) # marginally diff (p = 0.09)
kable(bmod8, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

## Ratio
bmod9<-aov(oligRatio ~ Year, data = zo.data) 
bmod9<-tidy(bmod9) # no diff
kable(bmod9, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod10<-aov(monoRatio ~ Year, data = zo.data) 
bmod10<-tidy(bmod10) # no diff
kable(bmod10, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
```
Significant differences: Larvae count, F count, brood mass, F mass, M mass is marginal

## T. stansburyi & Monodonts
### Plots
```{r}
## Count
boxplot(nTstanCells ~ Year, data = zo.data) # no diff, other than outlier
boxplot(nTstanCells ~ Year, data = zo.data.or) # no diff
boxplot(nParasite ~ Year, data = zo.data) # no diff

## Ratio
boxplot(tstanRatio ~ Year, data = zo.data) # similar
boxplot(tstanRatio ~ Year, data = zo.data.or) # similar
boxplot(monoRatio ~ Habitat, data = zo.data) # similar
```

### Analyses
```{r}
tmod1<-aov(nTstanCells ~ Year, data = zo.data)
tmod1<-tidy(tmod1) # no diff
kable(tmod1, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
tmod2<-aov(nTstanCells ~ Year, data = zo.data.or) 
tmod2<-tidy(tmod2) # no diff
kable(tmod2, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
tmod3<-aov(tstanRatio ~ Year, data = zo.data.or) 
tmod3<-tidy(tmod3) # no diff
kable(tmod3, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
```
No differences

## Climate variables
### Plots
```{r}
boxplot(HLI ~ Year, data=zo.data) # no diff
boxplot(Ann_ppt_mm ~ Year, data=zo.data) # no diff
boxplot(Ann_tmax_C ~ Year, data=zo.data) # no diff
boxplot(Ann_tmean_C ~ Year, data=zo.data) # no diff
boxplot(Ann_tmin_C ~ Year, data=zo.data) # no diff

boxplot(bi_ppt_mm ~ Year, data = zo.data) # diff
boxplot(bi_tmin_C ~ Year, data = zo.data) # diff
boxplot(bi_tmean_C ~ Year, data = zo.data) # diff
boxplot(bi_tmax_C ~ Year, data = zo.data) # diff
boxplot(qrt_ppt_mm ~ Year, data = zo.data) # diff
boxplot(qrt_tmin_C ~ Year, data = zo.data) # no diff
boxplot(qrt_tmean_C ~ Year, data = zo.data) # no diff
boxplot(qrt_tmax_C ~ Year, data = zo.data) # diff
boxplot(DD_accumulated ~ Year, data=zo.data) # diff
```
Precipitation and temperatures from Jan-Jun will probably be different. ADD is also MUCH higher in 2020 - boxes were left out longer and temperatures were higher in 2020

### Analyses
```{r}
cmod1<-aov(HLI ~ Year, data=zo.data) 
cmod1<-tidy(cmod1) # no diff
kable(cmod1, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod2<-aov(Ann_ppt_mm ~ Year, data=zo.data) 
cmod2<-tidy(cmod2) # no diff
kable(cmod2, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod3<-aov(Ann_tmax_C ~ Year, data=zo.data) 
cmod3<-tidy(cmod3) # no diff
kable(cmod3, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod4<-aov(Ann_tmean_C ~ Year, data=zo.data) 
cmod4<-tidy(cmod4) # no diff
kable(cmod4, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod5<-aov(Ann_tmin_C ~ Year, data=zo.data) 
cmod5<-tidy(cmod5) # no diff
kable(cmod5, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod6<-aov(bi_ppt_mm ~ Year, data = zo.data) 
cmod6<-tidy(cmod6) # diff
kable(cmod6, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod7<-aov(bi_tmin_C ~ Year, data = zo.data) 
cmod7<-tidy(cmod7) # diff
kable(cmod7, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod8<-aov(bi_tmean_C ~ Year, data = zo.data) 
cmod8<-tidy(cmod8) # diff
kable(cmod8, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod9<-aov(bi_tmax_C ~ Year, data = zo.data) 
cmod9<-tidy(cmod9) # diff
kable(cmod9, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod10<-aov(qrt_ppt_mm ~ Year, data = zo.data) 
cmod10<-tidy(cmod10) # diff
kable(cmod10, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod11<-aov(qrt_tmin_C ~ Year, data = zo.data) 
cmod11<-tidy(cmod11) # no diff
kable(cmod11, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod12<-aov(qrt_tmean_C ~ Year, data = zo.data) 
cmod12<-tidy(cmod12) # no diff
kable(cmod12, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod13<-aov(qrt_tmax_C ~ Year, data = zo.data) 
cmod13<-tidy(cmod13) # diff
kable(cmod13, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

cmod14<-aov(DD_accumulated ~ Year, data=zo.data) 
cmod14<-tidy(cmod14) # diff
kable(cmod14, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
```
Significant differences: bi_ppt_mm, bi_tmin_C, bi_tmean_C, bi_tmax_C, qrt_ppt_mm, qrt_tmax_C, and DD_accumulated

## Floral variables
### Plots
```{r}
boxplot(stemdiv ~ Year, data=zo.data) # higher in 2020
boxplot(stemrich ~ Year, data=zo.data) # higher in 2020
```

### Analysis
```{r}
fmod1<-aov(stemdiv ~ Year, data=zo.data) 
fmod1<-tidy(fmod1) # diff
kable(fmod1, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

fmod2<-aov(stemrich ~ Year, data=zo.data)
fmod2<-tidy(fmod2) # diff
kable(fmod2, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
```
Significant vars: stemdiv, and stemrich

## Trees
### Plots
```{r}
boxplot(meanTreeCC ~ Year, data=zo.data) # no diff
boxplot(Dead_BA_ha ~ Year, data=zo.data) # slight diff in median, probably not sig
boxplot(Live_BA_ha ~ Year, data=zo.data) # no diff
```

### Analysis
```{r}
tmod1<-aov(meanTreeCC ~ Year, data=zo.data) 
tmod1<-tidy(tmod1) # no diff
kable(tmod1, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

tmod2<-aov(Dead_BA_ha ~ Year, data=zo.data) 
tmod2<-tidy(tmod2) # no diff
kable(tmod2, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

tmod3<-aov(Live_BA_ha ~ Year, data=zo.data) 
tmod3<-tidy(tmod3) # no diff
kable(tmod3, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
```
No significant differences

## Landscape variables
### Plots
```{r}
boxplot(EVT_abund ~ Year, data=zo.data) # higher in 2020
boxplot(EVT_H ~ Year, data=zo.data) # higher in 2021
boxplot(EVT_J ~ Year, data=zo.data) # slightly higher in 2020
boxplot(Barren ~ Year, data=zo.data) # somewhat higher in 2021
boxplot(Coniferous_Forest ~ Year, data=zo.data) # no diff
boxplot(Crops ~ Year, data=zo.data) # no diff
boxplot(Deciduous_Forest ~ Year, data=zo.data) # no diff
boxplot(Rangeland ~ Year, data=zo.data) # no diff
boxplot(WUI ~ Year, data=zo.data) # higher in 2021
```
Probably some differences in EVT_abudn, EVT_H, EVT_J, Barren, and WUI

### Analysis
```{r}
lmod1<-aov(EVT_abund ~ Year, data=zo.data) 
lmod1<-tidy(lmod1) # diff
kable(lmod1, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

lmod2<-aov(EVT_H ~ Year, data=zo.data) 
lmod2<-tidy(lmod2) # marginally diff
kable(lmod2, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

lmod3<-aov(EVT_J ~ Year, data=zo.data) 
lmod3<-tidy(lmod3) # marginally diff
kable(lmod3, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

lmod4<-aov(Barren ~ Year, data=zo.data) 
lmod4<-tidy(lmod4) # marginally diff
kable(lmod4, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

lmod5<-aov(Coniferous_Forest ~ Year, data=zo.data) 
lmod5<-tidy(lmod5) # no diff
kable(lmod5, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

lmod6<-aov(Crops ~ Year, data=zo.data) 
lmod6<-tidy(lmod6) # no diff
kable(lmod6, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

lmod7<-aov(Deciduous_Forest ~ Year, data=zo.data) 
lmod7<-tidy(lmod7) # no diff
kable(lmod7, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

lmod8<-aov(Rangeland ~ Year, data=zo.data) 
lmod8<-tidy(lmod8) # no diff
kable(lmod8, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

lmod9<-aov(WUI ~ Year, data=zo.data) 
lmod9<-tidy(lmod9) # marginally diff
kable(lmod9, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
```
Significant vars: EVT_abund
Marginal sig vars: EVT_H, EVT_J, Barren, WUI

## Sum stats
For significantly different variables between years
### O. lignaria
Significant differences: Larvae count, F count, F mass, M mass is marginal
```{r}
zo.data %>% group_by(Year) %>% 
  summarise_at(c("nOligF", "nOligLarve", "avgOligFMass_mg", "avgOligMMass_mg"), 
                funs(mean=round(mean(.), 3), se=round(sd(.), 3))) %>% print.data.frame() 
```

### Climate
Significant differences: bi_ppt_mm, bi_tmin_C, bi_tmean_C, bi_tmax_C, and DD_accumulated
```{r}
# Field days between years
zo.data %>% group_by(Year) %>% 
  summarise_at(c("bi_ppt_mm", "bi_tmin_C", "bi_tmean_C", "bi_tmax_C", "FieldDay", "DD_accumulated"), 
                funs(mean=round(mean(.), 3), se=round(sd(.), 3))) %>% print.data.frame() 
```

### Floral
Significant vars: stemdiv, and stemrich
```{r}
zo.data %>% group_by(Year) %>% 
  summarise_at(c("stemdiv", "stemrich"), 
               funs(mean=round(mean(.), 3), se=round(sd(.), 3))) %>% print.data.frame() 
```

### Landscape
Significant vars: EVT_abund
Marginal sig vars: EVT_H, EVT_J, Barren, WUI
```{r}
zo.data %>% group_by(Year) %>% 
  summarise_at(c("EVT_abund", "EVT_H", "EVT_J", "Barren", "WUI"), 
               funs(mean=round(mean(.), 3), se=round(sd(.), 3))) %>% print.data.frame() 
```

# Differences by habitat
## Data prep
```{r}
habdata <- zo.data[,-c(3)]
summary(habdata)
```

## O. lignaria
### Plots
```{r}
## Count
boxplot(nTotProvised ~ Habitat, data = zo.data) # more in treated
boxplot(nOligGrowth ~ Habitat, data = zo.data) # no diff in medians
boxplot(nOligLarve ~ Habitat, data=zo.data) # slightly diff
boxplot(nOligF ~ Habitat, data = zo.data) # slightly diff
boxplot(nOligM ~ Habitat, data = zo.data) # no diff

## Mass
boxplot(avgOligMass_mg ~ Habitat, data = zo.data) # diff in variance
boxplot(avgOligFMass_mg ~ Habitat, data = zo.data) # diff variance
boxplot(avgOligMMass_mg ~ Habitat, data = zo.data) # diff variance

## Ratio
boxplot(oligRatio ~ Habitat, data = zo.data) # similar
```

#### Count distribution
```{r}
ggplot(habdata, aes(nOligGrowth, fill = Habitat)) + geom_histogram(binwidth = 5) + facet_grid(Habitat ~ 
    ., margins = TRUE)

with(habdata, tapply(nOligGrowth, Habitat, function(x) {
    sprintf("M (SD) = %1.2f (%1.2f)", mean(x), sd(x))
}))
```

### *Analysis*
Going with lm, but might not be best fit (normality?)
Note: anova might not be best fit, because data is not balanced
```{r}
x = colnames(habdata[,-c(2)]) # make list of response vars names

#y = sort(unique(unlist(lapply(habdata, names))))
#i = sapply(habdata, "[", y)

olig.hab.out <- unlist(lapply(1, function(n) combn(x, 1, FUN=function(row) paste0(paste0(row, collapse = "+"), " ~ Habitat"))))
olig.hab.out

#To have the regression coefficients
olig.hab.mods = bind_rows(lapply(olig.hab.out, function(frml) {
  a = tidy(lm(frml, na.action=na.exclude, data = habdata[(. != Inf & is.na(.) == FALSE)])) # different from na.exclude
  a$frml = frml
  return(a)
}))

# check modes
olig.hab.mod.chk = bind_rows(lapply(olig.hab.out, function(frml) {
  b = plot(lm(frml, data=olig.habdata))
  b$frml= frml
  return(b)
}))

head(olig.hab.mods)
```
This doesn't work - have to overcome NA's in y variables

### *Analyses -long way*
```{r}
## Count
bmod1<-aov(nTotProvised ~ Habitat, data = zo.data) 
bmod1<-tidy(bmod1) # no diff
kable(bmod1, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod2<-aov(nOligGrowth ~ Habitat, data = zo.data)
bmod2<-tidy(bmod2) # no diff
kable(bmod2, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod3<-aov(nOligLarve ~ Habitat, data=zo.data) 
bmod3<-tidy(bmod3) # very diff
kable(bmod3, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod4<-aov(nOligF ~ Habitat, data = zo.data) 
bmod4<-tidy(bmod4) # very diff
kable(bmod4, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod5<-aov(nOligM ~ Habitat, data = zo.data) 
bmod5<-tidy(bmod5) # no diff
kable(bmod5, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

## Mass
bmod6<-aov(avgOligMass_mg ~ Habitat, data = zo.data) 
bmod6<-tidy(bmod6) # diff
kable(bmod6, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod7<-aov(avgOligFMass_mg ~ Habitat, data = zo.data) 
bmod7<-tidy(bmod7) # diff
kable(bmod7, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod8<-aov(avgOligMMass_mg ~ Habitat, data = zo.data) 
bmod8<-tidy(bmod8) # marginally diff (p = 0.09)
kable(bmod8, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )

## Ratio
bmod9<-aov(oligRatio ~ Habitat, data = zo.data) 
bmod9<-tidy(bmod9) # no diff
kable(bmod9, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
bmod10<-aov(monoRatio ~ Habitat, data = zo.data) 
bmod10<-tidy(bmod10) # no diff
kable(bmod10, col.names = c("Term","Df", "Sum Sq", "Mean Sq", "*t-statistic*", "*P-value*"), digits = 3 )
```


# How to analyze data
https://drizopoulos.github.io/GLMMadaptive/articles/ZeroInflated_and_TwoPart_Models.html#two-parthurdle-poisson-mixed-effects-model

* Weight data
  ** is continuous data with lots of zeros, or semi-continuous data. Therefore, a *Two-Part Mixed Effects Model for Semi-Continuous Data* might be appropriate
    *** this is a two step model: 
      **** 1) uses a logistic regression to determine if the outcome is zero or not (dichotomous) 
      **** 2) if data is not zero, uses a standard linear mixed model 
  ** scratch that - it doesn't make since to include zeros for weight data, only count data. Therefore, just do a glm.
  *** Note: Do NOT want to use a Poisson dist (only useful for integer data)

* Count data
  ** is discrete data and we want to know the distribution. Commonly, a Poisson distribution (generalized linear model family) is used for count data.
  ** However, our data is very zero heavy 
  ** Can use a *Zero-inflated Poisson Mixed Effects Model*
    *** user has the option to leave **zi_random** set to **NULL**, in which case for the zero-part we have a logistic regression with only fixed effects and no random effects
  ** Another option is the *Two-Part/Hurdle Poisson Mixed Effects Model**
    *** similar to the semi-continuous data data (above) but for discrete data (i.e. counts) instead of continuous data
    *** 1) uses logistic regression to determine if data is zero or not
    *** 2) if value is positive, used a truncated zero-inflated Poisson mixed effects model

## Q1 
* T. stansburyi abundance by landscape, climate, and floral variables
  ** What variables to include?
  ** Predictor vars to include in stp-wise model: https://journals.lww.com/picp/fulltext/2017/08030/Common_pitfalls_in_statistical_analysis__Logistic.9.aspx
    1) What vars have a sig, univariate relationship (P < 0.1)
    2) Avoid highly correlated variables

## Q2 
* T. stansburyi presence depends on O.lig provisions - 
  ** Omit sites that don't have provisions
  ** Note: different than sites with no O.lig ratios
    *** 1 site in 2021 was completely parasitized by T. stan
* 2020
  ** ntstanCells
    *** is zero influenced (zi), has a HUGE outlier, and not normal
      **** log + 1 transformation seems to help, but is still zero heavy, then has a normal dist 
      **** ZINB or ZIP might also be required
        ***** Or can we just do a ZINB/ZIP without transforming?
  ** nstan ratio
    *** not zi, but is zero heavy and not normal
    *** transformations don't seem to help
      ***** glm with poisson dist?
* 2021
  ** ntstanCells
    *** not zi, but is zero heavy and not normal
    *** sqrt and log+1 transformations help
      **** lm/glm
  ** tstan ratio
    *** not zi, but is zero heavy and not normal
    *** sqrt transformation helps
      **** lm/glm

